\documentclass[12pt,letterpaper,final]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\renewcommand\qedsymbol{$\blacksquare$}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{pdfpages}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usetikzlibrary{automata,arrows}
\usepackage{mdframed}
\usepackage[
  backend=biber,
  style=authoryear,
  sorting=nyt          % sort by name, year, title
]{biblatex}
\addbibresource{references.bib}

% Based on template from COSC-417 at Towson University, originally authored by Marius Zimand

\begin{document}

\fbox{
  \vbox{
    \begin{flushleft}
      Jonathan Llovet \\  % authors' names
      EN.605.202.81 \\  %class
      2026-06-02 \\  % date
      Module 2 - Homework 2 \\ % ID
    \end{flushleft}
    \center{\Large{\textbf{Stacks - Data Structures}}}
    %\end{mdframed}
  } % end vbox
} % end fbox
\vline

Write pseudo-code for problems requiring code. Do not write Java, Python or C++. You are responsible for the appropriate level of detail.

\medskip

\noindent\textbf{Problem 1:}
Using stack operations: push, pop, peek and empty:

\begin{enumerate}
  \item Write a pseudo-code function which returns the value of the item at the bottom of the stack, leaving the stack unchanged. (Hint: use an auxiliary stack.)
  \item Write a pseudo-code function which returns the value of the third element from the bottom of the stack. The stack may be left changed.
\end{enumerate}

\pagebreak
\noindent\textbf{Answer:}

\begin{mdframed}
  \begin{verbatim}
function peekBottom(s Stack) {
  raise an error if s is null
  bottom = null
  if empty(s) {
    // s doesn't contain anything,
    // so we just return the null value in bottom.
    return bottom
  }
  auxStack = Stack() Unbounded Stack
  // Move items from s to auxStack until s is empty.
  while true {
    top = pop(s)
    push(auxStack, top)
    if empty(s) {
      bottom = top // store this to return
      break out of loop
    }
  }
  // Put things back the way they were.
  // Move items back from auxStack to s
  // until auxStack is empty.
  while true {
    top = pop(auxStack)
    push(s, top)
    if empty(auxStack) {
      break out of loop
    }
  }
  return bottom
}
\end{verbatim}
\end{mdframed}

\pagebreak
\begin{mdframed}
  \begin{verbatim}
function peekNthFromBottom(s Stack, n int) {
  raise an error if s is null
  validate that n is nonnegative
  bottom = null
  if empty(s) {
    // s doesn't contain anything,
    // so we just return the null value in bottom.
    return bottom
  }
  auxStack = Stack() // Unbounded Stack
  // Move items from s to auxStack until s is empty.
  while true {
    top = pop(s)
    push(auxStack, top)
    if empty(s) {
      break out of loop
    }
  }
  // Use a counter to track number of pops
  counter = 1
  datum = null
  // Count up to n, pushing intermediate items back onto s.
  while counter <= n {
    datum = pop(auxStack)
    push(s, datum) // push back onto s
    counter = counter + 1
  }
  // Finish putting things back the way they were.
  while true {
    top = pop(auxStack)
    push(s, top)
    if empty(auxStack) {
      break out of loop
    }
  }
  return datum
}
\end{verbatim}
\end{mdframed}

To get the antepenultimate item on the stack, call
\[
  \verb|peekNthFromBottom(s Stack, n int)|
\]
with a given stack \verb|s| and \verb|n = 3|:
\[
  \verb|peekNthFromBottom(s, 3)|
\]

\pagebreak
\noindent\textbf{Problem 2:}
Simulate the action of the algorithm for checking delimiters for each of these strings by using a stack and showing the contents of the stack at each point. Do not write an algorithm.

\noindent\textbf{Answer:}

\begin{enumerate}
  \item \{[A+B]-[(C-D)]
        \begin{center}
          \begin{tabular}{|c|l|l|}
            \hline
            \text{Input} & \text{Stack}                   & \text{Result}     \\ \hline
                         &                                &                   \\
            $\{$         & $ $                            & Push $\boxed{\{}$ \\ [10pt]
            $[$          & $\boxed{\{}$                   & Push $\boxed{[}$  \\ [10pt]
            $A$          & $\boxed{\{}\boxed{[}$          & Ignore $A$        \\ [10pt]
            $+$          & $\boxed{\{}\boxed{[}$          & Ignore $+$        \\ [10pt]
            $B$          & $\boxed{\{}\boxed{[}$          & Ignore $B$        \\ [10pt]
            $]$          & $\boxed{\{}\boxed{[}$          & Pop $\boxed{[}$   \\ [10pt]
            $-$          & $\boxed{\{}$                   & Ignore $-$        \\ [10pt]
            $[$          & $\boxed{\{}$                   & Push \boxed{[}    \\ [10pt]
            $($          & $\boxed{\{}\boxed{[}$          & Push \boxed{(}    \\ [10pt]
            $C$          & $\boxed{\{}\boxed{[}\boxed{(}$ & Ignore $C$        \\ [10pt]
            $-$          & $\boxed{\{}\boxed{[}\boxed{(}$ & Ignore $-$        \\ [10pt]
            $D$          & $\boxed{\{}\boxed{[}\boxed{(}$ & Ignore $D$        \\ [10pt]
            $)$          & $\boxed{\{}\boxed{[}\boxed{(}$ & Pop \boxed{(}     \\ [10pt]
            $]$          & $\boxed{\{}\boxed{[}$          & Pop \boxed{[}     \\ [10pt]
            $\ $         & $\boxed{\{}$                   & Reject            \\ [10pt]
            \hline
          \end{tabular}
        \end{center}
        The stack was not empty at the end of the input. So, the program would reject or return false to indicate that the pairings are not valid.
        \pagebreak
  \item ((H) * {([J+K])})
        \begin{center}
          \begin{tabular}{|c|l|l|}
            \hline
            \text{Input} & \text{Stack}                            & \text{Result}     \\ \hline
                         &                                         &                   \\
            $($          & $ $                                     & Push $\boxed{(}$  \\ [10pt]
            $($          & $\boxed{(}$                             & Push $\boxed{(}$  \\ [10pt]
            $H$          & $\boxed{(}\boxed{(}$                    & Ignore $H$        \\ [10pt]
            $)$          & $\boxed{(}\boxed{(}$                    & Pop $\boxed{(}$   \\ [10pt]
            $*$          & $\boxed{(}$                             & Ignore $*$        \\ [10pt]
            $\{$         & $\boxed{(}$                             & Push $\boxed{\{}$ \\ [10pt]
            $($          & $\boxed{(}\boxed{\{}$                   & Push $\boxed{(}$  \\ [10pt]
            $[$          & $\boxed{(}\boxed{\{}\boxed{(}$          & Push $\boxed{[}$  \\ [10pt]
            $J$          & $\boxed{(}\boxed{\{}\boxed{(}\boxed{[}$ & Ignore $J$        \\ [10pt]
            $+$          & $\boxed{(}\boxed{\{}\boxed{(}\boxed{[}$ & Ignore $+$        \\ [10pt]
            $K$          & $\boxed{(}\boxed{\{}\boxed{(}\boxed{[}$ & Ignore $K$        \\ [10pt]
            $]$          & $\boxed{(}\boxed{\{}\boxed{(}\boxed{[}$ & Pop $\boxed{[}$   \\ [10pt]
            $)$          & $\boxed{(}\boxed{\{}\boxed{(}$          & Pop $\boxed{(}$   \\ [10pt]
            $\}$         & $\boxed{(}\boxed{\{}$                   & Pop $\boxed{\{}$  \\ [10pt]
            $)$          & $\boxed{(}$                             & Pop $\boxed{(}$   \\ [10pt]
            $\ $         &                                         & Accept            \\ [10pt]

            \hline
          \end{tabular}
        \end{center}
        The stack was empty at the end of the input. So, the program would accept or return true to indicate that the pairings are valid.
\end{enumerate}

\pagebreak
\noindent\textbf{Problem 3:}
Write an algorithm to determine whether an input character string is of the form $xCy$, where $x$ is a string consisting only of the letters `A' and `B' and $y$ is the reverse of the $x$ (i.e.\ if $x$=``ABABBA'' then y must equal ``ABBABA'').
\begin{itemize}
  \item At each point you may read only the next character in the string, i.e.\ you must process the string on a left to right basis one character at a time
  \item You may not use string functions (ie reverse)
\end{itemize}

\noindent\textbf{Answer:}

This is a problem well suited for a pushdown automaton (PDA). A PDA is a simple model of computation that reads a single symbol at a time from input and can use a stack. In a recent class I took on the theory of computation where we read \cite{sipser2013}, we used these for decision problems about whether strings had a specified form like this one. (This could be handled by a context free grammar as well, but it is more straightforward to write an algorithm using a PDA in this context.)

The diagram below shows an algorithm for validating that a string has the specified format . Each node in the graph represents a state that the computation is in. The edges between the nodes are transitions between those states. For this PDA diagram, the transitions should be read in the following way: $a, b \to c$ indicates that the symbol $a$ is read from input, $b$ is popped from the stack, and $c$ is pushed onto the stack. The special value $\varepsilon$ indicates the empty string or no value. Additionally, per convention with such diagrams, I am using $\$$ as a special symbol to indicate the bottom of the stack. Popping $\$$ from the stack is equivalent to having \verb|empty(s)| return \verb|true|. Transitions between states that are marked with $\varepsilon$ as the input symbol can be made without reading a symbol from input, such as the transition from $q_0$ to $q_1$, which is just necessary plumbing to set the $\$$ value at the bottom of the stack. Otherwise, transitions can only occur when reading the corresponding symbol from input and popping the corresponding symbol from the stack. A transition from one state to another can be performed if and only if there is an explicitly labelled transition matching the current input symbol and the current value popped from the stack. Finally, if a state has a double circle, it is an "accepting" state, which means here that the string is in the format specified. We will only have one accepting state in this diagram, and for any string that is in the specified format, there will be a path from the starting state $q_0$ to that accepting state $q_3$. If there is a character of input or a combination of input and the value popped from the stack that is not mapped by a transition in the graph, then the string is not in the required format, and the PDA rejects the input string.

\begin{center}
  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=3cm]
    \node[initial, state]   (q0) {$q_0$};
    \node[state]            (q1) [right of=q0] {$q_1$};
    \node[state]            (q2) [above right of=q1] {$q_2$};
    \node[state, accepting] (q3) [right of=q2] {$q_3$};

    \path[->]
    (q0) edge              node {$\varepsilon,\, \varepsilon \to \$$}         (q1)
    (q1) edge [loop right] node [right] {$A,\, \varepsilon \to A$}            (q1)
    (q1) edge [loop below] node {$B,\, \varepsilon \to B$}                    (q1)
    (q1) edge  [bend left] node {$C,\, \varepsilon \to \varepsilon$}          (q2)
    (q2) edge [loop above] node [above right] {$A,\, A \to \varepsilon$}      (q2)
    (q2) edge [loop below] node [below right] {$B,\, B \to \varepsilon$}      (q2)
    (q2) edge              node {$\varepsilon,\, \$ \to \varepsilon$}         (q3)
    ;
  \end{tikzpicture}
\end{center}

Walking through this algorithm:
\begin{itemize}
  \item Start computation at $q_0$.
  \item Transition without input to $q_1$ and push $\$$ onto the stack to indicate the bottom (equivalent to initializing a stack in a program).
  \item For each input symbol that is an $A$ or $B$, read the symbol and push the same symbol onto the stack, but loop back to $q_1$. If anything other than $A$, $B$, or $C$ is read from the input, then reject the string.
  \item When an input symbol $C$ is encountered, transition to $q_2$ and do nothing to the stack.
  \item For the next input symbol, which must be $A$ or $B$, the value popped of the stack must match the input symbol. If anything besides $A$ or $B$ is read from input, reject the string. If the symbol popped from the stack does not match the symbol read from input, reject the string. Loop back to $q_2$ if $A$ or $B$ is read and the popped symbol matches. Repeat this until all of the input symbols are read.
  \item If all of the symbols have been read from input and the stack is empty, then transition from $q_2$ to $q_3$, which is the accepting state. During the transition, pop the $\$$ off the stack, which is equivalent to having \verb|empty(s)| return \verb|true|. Accept the string.
  \item If all of the symbols have been read from input and the stack is not empty, then the string does not match the specified format, and there will not be a transition defined in the graph. Reject the string.
\end{itemize}

An interesting edge case worth considering is when a string is in the format $xCy$, but $x$ and $y$ are both empty. In this case, the computation will run as follows:

\begin{itemize}
  \item transition from start to $q_o$
  \item transition from $q_0$ to $q_1$: $\varepsilon, \varepsilon \to \$$
  \item transition from $q_1$ to $q_2$: $C, \varepsilon \to \varepsilon$
  \item transition from $q_2$ to $q_3$: $\varepsilon, \$ \to \varepsilon$
  \item accept
\end{itemize}

\noindent\textbf{Problem 4:}
Consider a language that does not have arrays but does have stacks defined as a data type.
\begin{enumerate}
  \item That is, one can declare: `stack s' which defines push, pop, empty, and peek operations. Show how a one-dimensional array can be implemented by using these operations on two stacks.
  \item Show how you can insert into and read from such an array.
\end{enumerate}

\noindent\textbf{Answer:}

An array \texttt{C} can be implemented by using two stacks in concert. Let the two stacks be \texttt{A} and \texttt{B}. The elements of the array are contained in \texttt{A} to start, with the top of \texttt{A} equal to the first element in \texttt{C}, the bottom of \texttt{A} equal to the last element in \texttt{C}, and the intermediate elements of the stack \texttt{A} equal to the corresponding elements of \texttt{C} in order. That is, \texttt{C[0]} is equal to \texttt{pop(A)}, \texttt{C[1]} to the result of executing \texttt{pop(A)} twice, and so on, with \texttt{C[n]} equal to the result of executing \texttt{pop(A)} \texttt{n+1} times. The stack \texttt{B} is initially empty.

To describe the rest of the implementation of the array, we have to consider reading from and writing to the array.

To read from the array while it is in this initial configuration, we can pop from \texttt{A} and push to \texttt{B} until we reach the element in \texttt{A} we are interested in reading. Let there be a function \texttt{transfer(\alpha, \beta, n)} that accepts two stacks \texttt{\alpha} and \texttt{\beta}, along with an integer \texttt{n}. This function pops \texttt{n} elements from \texttt{\alpha} and pushes each element onto \texttt{\beta} as is it popped.

Let there be a function \texttt{read(\gamma, index)}, which takes an array \texttt{\gamma} and an integer \texttt{index}. The function \texttt{read} can use the function \texttt{transfer(\alpha, \beta, n)} as a helper function to iterate through the elements of \texttt{\gamma} until the desired element at \texttt{index} is reached. That element can be peeked or popped to read the element from the array. Once the element at \texttt{\gamma[index]} is read, the transfer function can be called again, but with \texttt{\alpha} and \texttt{\beta} reversed, i.e. \texttt{transfer(\beta, \alpha, n)}, which will reset the elements back to their original state.

Writing is similar. Use \texttt{transfer} to move elements from \texttt{\alpha} to \texttt{\beta}, until the desired location is reached, then push the desired element onto \texttt{\alpha}, then finally transfer elements back from \texttt{\beta} to \texttt{\alpha}. Removing an element is the same, but with a \texttt{pop} whose value is discareded, rather than transferred.

As a word of commentary: this implementation can be thought of as using a cursor, which is implicitly at the top of stack \texttt{A}. More efficient implementations could keep track of additional fields, such as the cursor index on the array \texttt{C}, so that the reset step done after reading and writing in the implementation above could be avoided in many cases. If the array needs performance similar to a regular array, that would be important to do. In the same vein of thinking, this implementation is similar to a doubly linked list, where we shift the node we are pointing at back and forth as we read. There are important differences though, for example that this implementation does not have something like a reference to the head of the list.

\pagebreak
\noindent\textbf{Problem 5:}
Design a method for keeping two stacks within a single linear array s[SPACESIZE] so that neither stack overflows until all of memory is used and an entire stack is never shifted to a different location within the array.
\begin{enumerate}
  \item Write methods push1, push2, pop1, and pop2 to manipulate the two stacks.
  \item Hint: the two stacks grow toward each other.
\end{enumerate}

\noindent\textbf{Answer:}

\noindent\textbf{Problem 6:}
Transform each of the following expressions to prefix and postfix expressions. \$ is exponentiation.

\noindent\textbf{Answer:}

I'm using the notation $\alpha \implies \beta$ to represent that the expression $\alpha$ can be transformed into the expression $\beta$ with one step.
\begin{enumerate}
  \item \texttt{(A+B)*(C\$(D-E)+F)-G}

        Postfix
        \begin{align*}
                   & \texttt{(A+B)*(C\$(D-E)+F)-G}                           \\
          \implies & \boxed{\texttt{(AB+)}}\texttt{*(C\$(D-E)+F)-G}          \\
          \implies & \boxed{\texttt{(AB+)(C\$(D-E)+F)*}}\texttt{-G}          \\
          \implies & \texttt{(AB+)(}\boxed{\texttt{C(D-E)\$}}\texttt{+F)*-G} \\
          \implies & \texttt{(AB+)(C(}\boxed{\texttt{DE-}}\texttt{)\$+F)*-G} \\
          \implies & \texttt{(AB+)(C(DE-)\$}\boxed{\texttt{F+}}\texttt{)*-G} \\
          \implies & \texttt{(AB+)(C(DE-)\$F+)*}\boxed{\texttt{G-}}          \\
          \implies & \texttt{AB+CDE-\$F+*G-}
        \end{align*}

        Prefix
        \begin{align*}
                   & \texttt{(A+B)*(C\$(D-E)+F)-G}                           \\
          \implies & \boxed{\texttt{(+AB)}}\texttt{*(C\$(D-E)+F)-G}          \\
          \implies & \texttt{(+AB)*(}\boxed{\texttt{\$C(D-E)}}\texttt{+F)-G} \\
          \implies & \texttt{(+AB)*(\$C}\boxed{\texttt{(-DE)}}\texttt{+F)-G} \\
          \implies & \texttt{(+AB)*}\boxed{\texttt{(+\$C(-DE)F)}}\texttt{-G} \\
          \implies & \boxed{\texttt{*(+AB)(+\$C(-DE)F)}}\texttt{-G}          \\
          \implies & \boxed{\texttt{-*(+AB)(+\$C(-DE)F)G}}                   \\
          \implies & \texttt{-*+AB+\$C-DEFG}
        \end{align*}

  \item \texttt{A+(((B-C)*(D-E)+F)/G)\$(H-J)}

        Postfix
        \begin{align*}
                   & \texttt{A+(((B-C)*(D-E)+F)/G)\$(H-J)}                           \\
          \implies & \texttt{A+(((}\boxed{\texttt{BC-}}\texttt{)*(D-E)+F)/G)\$(H-J)} \\
          \implies & \texttt{A+(((BC-)*(}\boxed{\texttt{DE-}}\texttt{)+F)/G)\$(H-J)} \\
          \implies & \texttt{A+(((BC-)*(DE-)+F)/G)\$(}\boxed{\texttt{HJ-}}\texttt{)} \\
          \implies & \texttt{A+((}\boxed{\texttt{(BC-)(DE-)*}}\texttt{+F)/G)\$(HJ-)} \\
          \implies & \texttt{A+(((BC-)(DE-)*}\boxed{\texttt{F+}}\texttt{)/G)\$(HJ-)} \\
          \implies & \texttt{A+(((BC-)(DE-)*F+)}\boxed{\texttt{G/}}\texttt{)\$(HJ-)} \\
          \implies & \texttt{A+(((BC-)(DE-)*F+)G/)}\boxed{\texttt{(HJ-)\$}}          \\
          \implies & \boxed{\texttt{A(((BC-)(DE-)*F+)G/)(HJ-)\$+}}                   \\
          \implies & \texttt{ABC-DE-*F+G/HJ-\$+}                                     \\
        \end{align*}

        Prefix
        \begin{align*}
                   & \texttt{A+(((B-C)*(D-E)+F)/G)\$(H-J)}                           \\
          \implies & \texttt{A+(((}\boxed{\texttt{-BC}}\texttt{)*(D-E)+F)/G)\$(H-J)} \\
          \implies & \texttt{A+(((-BC)*(}\boxed{\texttt{-DE}}\texttt{)+F)/G)\$(H-J)} \\
          \implies & \texttt{A+(((-BC)*(-DE)+F)/G)\$(}\boxed{\texttt{-HJ}}\texttt{)} \\
          \implies & \texttt{A+((}\boxed{\texttt{*(-BC)(-DE)}}\texttt{+F)/G)\$(-HJ)} \\
          \implies & \texttt{A+((}\boxed{\texttt{+*(-BC)(-DE)F)}}\texttt{/G)\$(-HJ)} \\
          \implies & \texttt{A+(}\boxed{\texttt{/(+*(-BC)(-DE)F)G}}\texttt{)\$(-HJ)} \\
          \implies & \texttt{A+}\boxed{\texttt{\$(/(+*(-BC)(-DE)F)G)(-HJ)}}          \\
          \implies & \boxed{\texttt{+A}}\texttt{\$(/(+*(-BC)(-DE)F)G)(-HJ)}          \\
          \implies & \texttt{+A\$/+*-BC-DEFG-HJ}                                     \\
        \end{align*}

\end{enumerate}

\pagebreak
\noindent\textbf{Problem 7:}
Transform each of the following expressions to infix expressions.

\noindent\textbf{Answer:}

I start by grouping with parentheses according to where there will be two operands for the operation.
Then I convert each operation to infix notation.
As above, I'm using the notation $\alpha \implies \beta$ to represent that the expression $\alpha$ can be transformed into the expression $\beta$ with one step.

\begin{enumerate}
  \item \texttt{++A-*\$BCD/+EF*GHI}
        \begin{align*}
                   & \texttt{++A-*\$BCD/+EF*GHI}                                         \\
          \implies & \texttt{+(+A(-(*(\$BC)D)(/(+EF)(*GH))))I}                           \\
          \implies & \boxed{\texttt{(+A(-(*(\$BC)D)(/(+EF)(*GH))))+I}}                   \\
          \implies & \texttt{(}\boxed{\texttt{A+}}\texttt{(-(*(\$BC)D)(/(+EF)(*GH))))+I} \\
          \implies & \texttt{(A+(}\boxed{\texttt{(*(\$BC)D)-(/(+EF)(*GH))}}\texttt{))+I} \\
          \implies & \texttt{(A+((*}\boxed{\texttt{(B\$C)}}\texttt{D)-(/(+EF)(*GH))))+I} \\
          \implies & \texttt{(A+(}\boxed{\texttt{((B\$C)*D)}}\texttt{-(/(+EF)(*GH))))+I} \\
          \implies & \texttt{(A+(((B\$C)*D)-}\boxed{\texttt{((+EF)/(*GH))}}\texttt{))+I} \\
          \implies & \texttt{(A+(((B\$C)*D)-(}\boxed{\texttt{(E+F)}}\texttt{/(*GH))))+I} \\
          \implies & \texttt{(A+(((B\$C)*D)-((E+F)/}\boxed{\texttt{(G*H)}}\texttt{)))+I} \\
          \implies & \texttt{(A+(((B\$C)*D)-((E+F)/(G*H)))+I}
        \end{align*}

  \item \texttt{ABCDE-+\$*EF*-}
        \begin{align*}
                   & \texttt{ABCDE-+\$*EF*-}                                     \\
          \implies & \texttt{(A(B(C(DE-)+)\$)*)(EF*)-}                           \\
          \implies & \boxed{\texttt{(A(B(C(DE-)+)\$)*)-(EF*)}}                   \\
          \implies & \texttt{(A(B(C(}\boxed{\texttt{D-E}}\texttt{)+)\$)*)-(EF*)} \\
          \implies & \texttt{(A(B(C(D-E)+)\$)*)-(}\boxed{\texttt{E*F}}\texttt{)} \\
          \implies & \texttt{(A(B(}\boxed{\texttt{C+(D-E)}}\texttt{)\$)*)-(E*F)} \\
          \implies & \texttt{(A(}\boxed{\texttt{B\$(C+(D-E))}}\texttt{)*)-(E*F)} \\
          \implies & \texttt{(}\boxed{\texttt{A*(B\$(C+(D-E)))}}\texttt{)-(E*F)} \\
          \implies & \texttt{(A*(B\$(C+(D-E))))-(E*F)}
        \end{align*}
\end{enumerate}


\noindent\textbf{Problem 8:}
Write a prefix pseudo-code function to accept an infix string and create the prefix form of that string, assuming that the string is read from right to left and that the prefix string is created from right to left. Handle variables, +,-,/,*,\$,(,).

\noindent\textbf{Answer:}


\pagebreak

\section*{Source Code}
The full source code, including LaTeX, tooling, other artifacts, and git history is available on my Github repository for the class here:
\begin{center}
  \url{https://github.com/jllovet/jhu-en-605-202-su26-data-structures}
\end{center}

\printbibliography

\end{document}