\documentclass[12pt,letterpaper,final]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\renewcommand\qedsymbol{$\blacksquare$}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{pdfpages}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usetikzlibrary{automata,arrows}
\usepackage[
  backend=biber,
  style=authoryear,
  sorting=nyt          % sort by name, year, title
]{biblatex}
\addbibresource{references.bib}

% Based on template from COSC-417 at Towson University, originally authored by Marius Zimand

\begin{document}

\fbox{
  \vbox{
    \begin{flushleft}
      Jonathan Llovet \\  % authors' names
      EN.605.202.81 \\  %class
      2026-05-26 \\  % date
      Module 1 - Homework 1 \\ % ID
    \end{flushleft}
    \center{\Large{\textbf{Complexity and ADTs - Data Structures}}}
    %\end{mdframed}
  } % end vbox
} % end fbox
\vline


\noindent\textbf{Problem 1:}  Show how nonnegative integers can be represented in an imaginary ternary computer using trits $(0,1,2)$ instead of bits $(0,1)$. Why don't we do things this way?

\noindent\textbf{Answer:}
Nonnegative integers can be represented in ternary in the same way that they are represented in other bases. Decimal, binary, ternary, octal, hexadecimal, and other similar number systems are all position-based number systems, where each position represents a power of the base of the number system. For a base $b$, a number is represented by a sequence of $n+1$ numbers in the format
\[a_{n} \cdot b^{n},a_{n-1} \cdot b^{n-1},\ldots,a_{1} \cdot b^{1},a_{0} \cdot b^{0},\]
where \[a \in \mathbb{Z},\ 0 \le a < b.\]
The number represented by this sequence is the sum of the elements of the sequence. For example, in decimal, we are working with base-$10$ numbers. Setting $b=10$, we can represent a number in base 10 as:
\begin{align*}
      & a_{n} \cdot b^{n},a_{n-1} \cdot b^{n-1},\ldots,a_{1} \cdot b^{1},a_{0} \cdot b^{0}     \\
  =\  & a_{n} \cdot 10^{n},a_{n-1} \cdot 10^{n-1},\ldots,a_{1} \cdot 10^{1},a_{0} \cdot 10^{0} \\
\end{align*}
The number itself represented by this sequence is usually written without the commas and without the base explicitly noted. If there is ambiguity in context about the base being used, it is shown with a single subscript at the end of the number. Hence, the expression above is simplified as follows:

\begin{align*}
      & a_{n} \cdot 10^{n},a_{n-1} \cdot 10^{n-1},\ldots,a_{1} \cdot 10^{1},a_{0} \cdot 10^{0} \\
  =\  & a_{n},a_{n-1},\ldots,a_{1},a_{0}                                                       \\
\end{align*}

Hence the familiar decimal notation: $1234$, which can alternatively be written $1234_{10}$. Binary and ternary follow the same scheme.

Binary:
\begin{align*}
      & a_{n} \cdot 2^{n},a_{n-1} \cdot 2^{n-1},\ldots,a_{1} \cdot 2^{1},a_{0} \cdot 2^{0} \\
  =\  & a_{n},a_{n-1},\ldots,a_{1},a_{0}                                                   \\
\end{align*}

Ternary:
\begin{align*}
      & a_{n} \cdot 3^{n},a_{n-1} \cdot 3^{n-1},\ldots,a_{1} \cdot 3^{1},a_{0} \cdot 3^{0} \\
  =\  & a_{n},a_{n-1},\ldots,a_{1},a_{0}                                                   \\
\end{align*}

Consider the first few non-negative integers represented in a few of the number systems we frequently encounter in computer science, along with ternary:

\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Decimal & Binary & Ternary & Octal & Hex    \\
    \hline
    $b=10$  & $b=2$  & $b=3$   & $b=8$ & $b=16$ \\
    \hline
    $0$     & $0$    & $0$     & $0$   & $0$    \\
    $1$     & $1$    & $1$     & $1$   & $1$    \\
    $2$     & $10$   & $2$     & $2$   & $2$    \\
    $3$     & $11$   & $10$    & $3$   & $3$    \\
    $4$     & $100$  & $11$    & $4$   & $4$    \\
    $5$     & $101$  & $12$    & $5$   & $5$    \\
    $6$     & $110$  & $20$    & $6$   & $6$    \\
    $7$     & $111$  & $21$    & $7$   & $7$    \\
    $8$     & $1000$ & $22$    & $8$   & $8$    \\
    $9$     & $1001$ & $100$   & $10$  & $9$    \\
    $10$    & $1010$ & $101$   & $11$  & $A$    \\
    $11$    & $1011$ & $102$   & $12$  & $B$    \\
    $12$    & $1100$ & $110$   & $13$  & $C$    \\
    $13$    & $1101$ & $111$   & $14$  & $D$    \\
    $14$    & $1110$ & $112$   & $15$  & $E$    \\
    $15$    & $1111$ & $120$   & $16$  & $F$    \\
    \hline
  \end{tabular}
\end{center}
\pagebreak
Numbers in binary and ternary might be represented in the following way to disambiguate them from their decimal equivalents.

\[
  10_2 \ne 10_{10} \ne 10_3
\]

From the preceding, we can see that these numerical representations to each other mathematically, but each has slightly different properties. Some of the bases are easily convertible to each other, for example when one base is an integer power of another (i.e., $log_{b_1}b_2 \in \mathbb{Z}^+$), in which case conversion between bases can be done by grouping digits. This is the easiest strategy when converting between binary and hexadecimal, for example. Such conversion is possible but requires more steps between ternary and the common bases used in computer science, such as octal and hexadecimal. It might be objected that this obstacle is a red herring, since there are other bases that have an analogous position to ternary as octal and hexadecimal have to binary (such as base 9 and 27). However, there are other disadvantages to ternary as a number system for classical computing. What is such a disadvantage? In a word, ternary introduces complexity to physical implementations of classical computers, and binary can be used to represent integers (and other numbers) with less physical complexity.

The larger the base, the more numbers can be expressed with a single position. This makes reading the numbers more and less familiar with respect to our default of base 10 from everyday life. However, more information is required to represent numbers of higher bases. For each position in a number's representation, we must choose among $b$ options for the value of $a$. In the case of binary, when constructing a number, we have to make one decision with only two options, viz. $(0,1)$. For ternary, we must make a decision with three options, viz. $(0,1,2)$. This has implications for physical implementations in hardware, since we would need to have a mechanism that could unambiguously be in one of three distinct states to represent each trit, as opposed to the two distinct states to represent bits in binary. It might be physically possible to create a system where each component has three distinct states, but this would impose additional complexity on the physical implementation. As shown above, integers can be converted among any pair of bases. So, we are incentivized to implement this using the simplest option available. This option is binary, because there are only two states that need to be distinguished from each other. What's more, by using binary as the implementation for our computer systems, we are able to take advantage of the simple mapping between bits and boolean values. This mapping means that the two states required in physical representations of binary numbers can also be used for creating digital logic circuits.

\bigskip

\noindent\textbf{Problem 2:}  If each element of an array $RM$, with 10 rows and 20 columns, stored in row major order, takes four bytes of space, where the first element of $RM$ starts at 100, what is the address of $RM[5][3]$ and $RM[9][19]$?

\noindent\textbf{Answer:}

When indexing matrices in row-major order to lay out the data in memory as a sequence of contiguous addresses, we iterate through all of the entries in a row before going on to the next row. \textcite{byteology2014array}

Here we are using 0-based indexing and indicating rows, then columns. Since we are indexing from $0$, $RM[5][3]$ is the cell in the 6th row, 4rd column. Assuming that the machine is byte addressable, since each element of the array consumes four byes, there are four bytes consumed per cell in the matrix.

Using 100 as the base address, we can see a few representative entries in the table below.
\begin{center}
  \begin{tabular}{|c|c|c|}
    \hline
    $100 + (20 \times 0 \times 4) + (0 \times 4)$ = 100 & $100 + (20 \times 0 \times 4) + (1 \times 4) = 104$ & $\ldots$ \\
    $100 + (20 \times 1 \times 4) + (0 \times 4) = 180$ & $100 + (20 \times 1 \times 4) + (1 \times 4) = 184$ & $\ldots$ \\
    $\ldots$                                            & $\ldots$                                            & $\ldots$ \\
    $\ldots$                                            & $\ldots$                                            & $\ldots$ \\
    \hline
  \end{tabular}
\end{center}

The address for a specific location in the table $RM[i][j]$, where $i$ is the row and $j$ is the column, is given by
\[
  100 + x_i + x_j
\]
where the row-based offset is $x_i = i \times 20 \times 4$, and the column-based offset is $x_j = j \times 4$. From the preceding, the address can be calculated by the function below, which takes the two values $i, j$ as arguments.

\[f(i,j) = 100 + 80i + 4j\]

For the entries we are interested in:

Address of $RM[5][3]$:
\begin{align*}
  f(5,3) & = 100 + 80(5) + 4(3) \\
         & = 100 + 400 + 12     \\
         & = 512                \\
\end{align*}

Address of $RM[9][19]$:
\begin{align*}
  f(9,19) & = 100 + 80(9) + 4(19) \\
          & = 100 + 720 + 76      \\
          & = 896                 \\
\end{align*}

\bigskip
\noindent\textbf{Problem 3:}  A lower triangular matrix is an nxn array in which has $a[i][j] == 0$ if $i<j$. What is the maximum number of non zero elements? How can they be stored in memory sequentially? Find a formula $k = f(i,j)$ to store location $a[i][j]$ in $k$ (you only want to store the nonzero elements). Do not write code. Code would work if you were manually converting the matrix from one form to the other all at once, but you need the formula to convert otherwise.

In questions 3 and 4, we suggest indexing the array from one, since matrices are normally indexed from (1,1). After you find your formula, then you can adjust it to index the array from zero.

\noindent\textbf{Answer:}

To build an intuition for a lower triangular matrix, let's look at a few examples. We will set the values of each entry to 0 when $i < j$, otherwise we will set them to 1.
\[
  A_1 =
  \begin{bmatrix}
    a_{11}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1
  \end{bmatrix}
\]

\[
  A_2 =
  \begin{bmatrix}
    a_{11} & a_{12} \\
    a_{21} & a_{22}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 0 \\
    1 & 1
  \end{bmatrix}
\]

\[
  A_3 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} \\
    a_{21} & a_{22} & a_{23} \\
    a_{31} & a_{32} & a_{33}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 0 & 0 \\
    1 & 1 & 0 \\
    1 & 1 & 1
  \end{bmatrix}
\]

\[
  A_4 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} & a_{14} \\
    a_{21} & a_{22} & a_{23} & a_{24} \\
    a_{31} & a_{32} & a_{33} & a_{34} \\
    a_{41} & a_{42} & a_{43} & a_{44}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 0 & 0 & 0 \\
    1 & 1 & 0 & 0 \\
    1 & 1 & 1 & 0 \\
    1 & 1 & 1 & 1
  \end{bmatrix}
\]

\[
  A_5 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} & a_{14} & a_{15} \\
    a_{21} & a_{22} & a_{23} & a_{24} & a_{25} \\
    a_{31} & a_{32} & a_{33} & a_{34} & a_{35} \\
    a_{41} & a_{42} & a_{43} & a_{44} & a_{45} \\
    a_{51} & a_{52} & a_{53} & a_{54} & a_{55} \\
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 0 & 0 & 0 & 0 \\
    1 & 1 & 0 & 0 & 0 \\
    1 & 1 & 1 & 0 & 0 \\
    1 & 1 & 1 & 1 & 0 \\
    1 & 1 & 1 & 1 & 1
  \end{bmatrix}
\]

\begin{enumerate}
  \item What is the maximum number of non-zero elements?

        Here's one pattern that is emerging from the preceding $n \times n$ matrices: There are $n$ 1's in the nth row, $n-1$ of them in the $n-1$th row, and so on. Considering an entire matrix, the number of 1's in each is $\sum_{i=1}^{n} i$, which is the sum of the first $n$ positive integers. The maximum number of non-zero elements in a lower-triangular $n \times n$ matrix is therefore given by the function
        \begin{equation}
          g(n) = \frac{n(n+1)}{2} \label{eq:sum_of_first_n_ints}
        \end{equation}
  \item How can they be stored in memory sequentially?

        We can use the constraints of the problem to inform the layout of these non-zero elements in memory. For each row of the matrix, there are at most $n$ non-zero elements in that row. Because there are $n$ rows in the matrix, we can store 1 element, then 2 elements, 3 elements, and so on, up to $n$ elements. By arranging the elements this way, we can enumerate the non-zero elements of the matrix in order. Mathematically speaking, what we get here is a bijection between the natural numbers and these elements, which means that we can provide a unique index as a label to each of the non-zero elements of the matrix. Put another way, this enumeration provides us a way to place all of those elements in a list.
  \item Find a formula $k = f(i,j)$ to store location $a[i][j]$ in $k$ (you only want to store the nonzero elements).

        To find the $k$th element of the array, we can use the formula that we were using previously to our advantage. The function $g(n)$ \eqref{eq:sum_of_first_n_ints} above gives use the sum of the first $n$ integers. As we stated in the previous subsection, that is the number of elements in the first $n$ rows. Based on our scheme above for enumerating the non-zero elements of matrix in row-major order, this sum can provide us the initial part of the offset from the beginning of the array $a$ by evaluating $g(i-1)$, where $i$ is the value that was provided as an argument to $f(i, j)$.

        This initial offset from the beginning of $a$ indicates, more precisely, the location of the beginning of the non-zero elements of row $i$ in the matrix. For example, if we are concerned with elements of the 4th row of the matrix, $i = 4$. Evaluating $g(i-1)$, we have $\frac{i(i-1)}{2} = \frac{4 \cdot 3}{2}$, which returns $6$. This is the index immediately preceding the elements of row 4.

        The second part of the offset is provided by the value of $j$, which indicates which element of the row $i$ we want. In particular, if we consider $i = 4$ and $j = 2$, we will need to offset two more elements from the 6th element in the array, in order to get to the desired element in the 4th row, 2nd column. Putting this together, the index $k$ of a particular element in the array $a$ described is above is given by the function

        \begin{align*}
          k & = \frac{(i - 1)((i - 1) + 1)}{2} + j \\
            & = \frac{(i - 1)i}{2} + j             \\
            & = \frac{i^2 - i}{2} + j
        \end{align*}

        Which can be expressed as the function

        \begin{equation}
          f(i,j) = \frac{i^2 - i}{2} + j
        \end{equation}

        In the preceding, the analysis was according to 1-based indexing. In a program that uses 0-based indexing for arrays, the values provided by the caller of the function are $i-1$ and $j-1$. To translate this to the functions that we just developed for determining the position in the array $a$, we must add 1 to $i$ and to $j$.

        In particular, the function we would need in a program is $f'(i,j)$, described below. (Note however, that in a real implementation, we would need to include input validation that enforced the constraints of our construction. For example, the input should only support indexing where $i \ge j$.)

        Consider the illustration using the $f(i,j)$ for a 1-based matrix $B_1$, where we have labelled the entries in the order they would be in memory, starting from 1.

        \[
          B_1 =
          \begin{bmatrix}
            b_{11}         & b_{12} & b_{13} & b_{14} & b_{15} \\
            \boxed{b_{21}} & b_{22} & b_{23} & b_{24} & b_{25} \\
            b_{31}         & b_{32} & b_{33} & b_{34} & b_{35} \\
            b_{41}         & b_{42} & b_{43} & b_{44} & b_{45} \\
            b_{51}         & b_{52} & b_{53} & b_{54} & b_{55} \\
          \end{bmatrix}
          =
          \begin{bmatrix}
            1         & 0  & 0  & 0  & 0  \\
            \boxed{2} & 3  & 0  & 0  & 0  \\
            4         & 5  & 6  & 0  & 0  \\
            7         & 8  & 9  & 10 & 0  \\
            11        & 12 & 13 & 14 & 15
          \end{bmatrix}
        \]

        We want to access the boxed value in our array $a$ by using the offset provided by $f(i,j)$. Let $i=2,j=1$.

        \begin{align*}
          f_{1-based}(i,j) & = \frac{i^2 - i}{2} + j \\
                           & = \frac{2^2 - 2}{2} + 1 \\
                           & = \frac{2}{2} + 1       \\
                           & = 2
        \end{align*}
        This is the value that we desired. Now, to adapt this to the 0-based implementation, we need to adjust our function. We will access the same location, but with the indices shifted to 0-based indexing.

        \[
          B_0 =
          \begin{bmatrix}
            b_{00}         & b_{01} & b_{02} & b_{03} & b_{04} \\
            \boxed{b_{10}} & b_{11} & b_{12} & b_{13} & b_{14} \\
            b_{20}         & b_{21} & b_{22} & b_{23} & b_{24} \\
            b_{30}         & b_{31} & b_{32} & b_{33} & b_{34} \\
            b_{40}         & b_{41} & b_{42} & b_{43} & b_{44} \\
          \end{bmatrix}
          =
          \begin{bmatrix}
            1         & 0  & 0  & 0  & 0  \\
            \boxed{2} & 3  & 0  & 0  & 0  \\
            4         & 5  & 6  & 0  & 0  \\
            7         & 8  & 9  & 10 & 0  \\
            11        & 12 & 13 & 14 & 15
          \end{bmatrix}
        \]
        For this, if the user calls the function with the arguments $f(1,0)$, we can access the appropriate element. Consider the following adaptation of the function $f(i,j)$ to be used with a 0-based matrix or multidimensional array to access an arbitrary element when the non-zero elements are arranged into a single linear array.
        \begin{align*}
          f'_{0-based}(i,j) & = \frac{(i+1)^2 - (i+1)}{2} + j + 1      \\
                            & = \frac{i^2 + 2i + 1 - i - 1}{2} + j + 1 \\
                            & = \frac{i^2 + i}{2} + j + 1
        \end{align*}

        Let $i = 1, j = 0$. Evaluating $f(i,j)$:

        \begin{align*}
          f'_{0-based}(1,0) & = \frac{i^2 + i}{2} + j + 1 \\
                            & = \frac{1^2 + 1}{2} + 0 + 1 \\
                            & = \frac{2}{2} + 1           \\
                            & = 2
        \end{align*}

        This is also the value desired.

\end{enumerate}

\bigskip
\noindent\textbf{Problem 4:}  A tridiagonal matrix is an $n \times n$ array in which has $a[i][j] == 0$ if $|i-j| > 1$. What is the maximum number of non zero elements? How can they be stored in memory sequentially? Find a formula $k = f(i,j)$ to store location $a[i][j]$ in $k$, when $|i-j| \le 1$ (you only want to store the nonzero elements). Do not write code. Code would work if you were manually converting the matrix from one form to the other all at once, but you need the formula to convert otherwise.

\noindent\textbf{Answer:}

To build an intuition for a tridiangular matrix, let's look at a few examples. We will set the values of each entry to 0 when $|i - j| > 1$, otherwise we will set them to 1.

\[
  A_1 =
  \begin{bmatrix}
    a_{11}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1
  \end{bmatrix}
\]

\[
  A_2 =
  \begin{bmatrix}
    a_{11} & a_{12} \\
    a_{21} & a_{22}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 1 \\
    1 & 1
  \end{bmatrix}
\]

\[
  A_3 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} \\
    a_{21} & a_{22} & a_{23} \\
    a_{31} & a_{32} & a_{33}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 1 & 0 \\
    1 & 1 & 1 \\
    0 & 1 & 1
  \end{bmatrix}
\]

\[
  A_4 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} & a_{14} \\
    a_{21} & a_{22} & a_{23} & a_{24} \\
    a_{31} & a_{32} & a_{33} & a_{34} \\
    a_{41} & a_{42} & a_{43} & a_{44}
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 1 & 0 & 0 \\
    1 & 1 & 1 & 0 \\
    0 & 1 & 1 & 1 \\
    0 & 0 & 1 & 1
  \end{bmatrix}
\]

\[
  A_5 =
  \begin{bmatrix}
    a_{11} & a_{12} & a_{13} & a_{14} & a_{15} \\
    a_{21} & a_{22} & a_{23} & a_{24} & a_{25} \\
    a_{31} & a_{32} & a_{33} & a_{34} & a_{35} \\
    a_{41} & a_{42} & a_{43} & a_{44} & a_{45} \\
    a_{51} & a_{52} & a_{53} & a_{54} & a_{55} \\
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & 1 & 0 & 0 & 0 \\
    1 & 1 & 1 & 0 & 0 \\
    0 & 1 & 1 & 1 & 0 \\
    0 & 0 & 1 & 1 & 1 \\
    0 & 0 & 0 & 1 & 1
  \end{bmatrix}
\]

A pattern that becomes evident from the examples above: There are three diagonals that are non-zero values (whence this type of matrix receives its name). In the cases where $n > 1$, the first and last rows of the matrix only have two non-zero values, while each of the intermediate rows have up to three non-zero values. This provides as means for finding the upper bound to the number of non-zero elements in the matrix and a pattern that can be used to construct an arrangement of the elements in memory.

\begin{enumerate}
  \item What is the maximum number of non-zero elements?

        For these $n \times n$ matrices, there are at most two non-zero elements in the first and last rows. (This is because the diagonals reach the upper left and bottom right corners of the matrices here, and consequently there is not sufficient space to have more non-zero elements.) The intermediate rows each have at most 3 non-zero elements.

        We may need to be sensitive to the possible boundary case $n = 1$. In this case, there is only a single element in the $1 \times 1$ matrix, and it is non-zero. In the case that $n \ge 2$, we have $2 \times 2 + 3(n-2)$ non-zero elements. Simplifying this, we have $4 + 3n - 6 = 3n - 2$ non-zero elements. However, if we evaluate $3n -2$ with $n = 1$, we find that this expression yields the correct result as well ($3(1) - 2 = 1$). Hence, we do not need to handle the case $n=1$ as a special instance.

        Therefore, the maximum number of non-zero elements in a tridiangular matrix is given accordingly by the function
        \begin{equation}
          g(n) = 3n - 2
        \end{equation}
  \item How can they be stored in memory sequentially?
        Similar to the case in the previous question, there is an enumeration of the elements, which means that we can uniquely associate the non-zero elements with integer indices of an array. We can arrange the non-zero elements in row-major order, and because we know how many non-zero elements are in each row of the matrix based on the size of $n$, we can create a function that gives us the index mentioned above. In the case $n=1$, we only need to store a single element, and it will have index 0. In the case $n=2$, we will store the two elements of the first row, then the two elements of the second row. In cases where $n \ge 3$, we can store the first two elements from row one, then each three elements of the intermediate rows, then the final two elements from the last row.
  \item Find a formula $k = f(i,j)$ to store location $a[i][j]$ in $k$, when $|i - j| \le 1$.

        As we did in the previous problem, we need to calculate the offset from start of the array $a$ based on the value of $i$ and what we know about the structure of the matrix.

        Assume the matrix uses 1-based indexing. If $i = 1$, then we are near the beginning of the array, and we can use $j$ as the position of the element directly. If $i > 1$, then we have to account for a column offset in addition to our row offset. If $i = 2$, then the first non-zero value is at $j = 1$. If $i = 3$, then the first non-zero value is at $j = 2$. In general, the first non-zero value will be at $j = i - 1$. Accordingly, in order to specify the correct location in the array $a$ that the non-zero values from the matrix have been mapped to, we need to find the initial offset and then adjust the value of $j$. How can we do this? Take for example the case $f(3,2)$ in the matrix $B_1$ below.

        \[
          B_1 =
          \begin{bmatrix}
            b_{11} & b_{12}         & b_{13} & b_{14} & b_{15} \\
            b_{21} & b_{22}         & b_{23} & b_{24} & b_{25} \\
            b_{31} & \boxed{b_{32}} & b_{33} & b_{34} & b_{35} \\
            b_{41} & b_{42}         & b_{43} & b_{44} & b_{45} \\
            b_{51} & b_{52}         & b_{53} & b_{54} & b_{55} \\
          \end{bmatrix}
          =
          \begin{bmatrix}
            1 & 1         & 0 & 0 & 0 \\
            1 & 1         & 1 & 0 & 0 \\
            0 & \boxed{1} & 1 & 1 & 0 \\
            0 & 0         & 1 & 1 & 1 \\
            0 & 0         & 0 & 1 & 1
          \end{bmatrix}
        \]

        The function we are designing should place us at $b_{32}$ by specifying $i=3$. An index of $j=2$ will not work here, since that will move us to $b_{33}$. To adjust the offset to account for how the beginning of the sequence of non-zero values is already offset, we can take advantage of what we noted above. To specify that we want the element at index $1$ of the subsequence of non-zero values in row $3$, we can rearrange the expression from above,

        \begin{align*}
          j         & = i - 1 \\
          j - i + 1 & = 0
        \end{align*}

        This results in a shift of 0 positions, leaving us at position $b_{32}$.

        Coming back to our row-based offset, we must note that if we are in the last row and $1 < i = n$, then we only have two elements in the subsequence. That should not impact our formula here, but we should note that in a real implementation we would need to provide validation to ensure that we did not index out of bounds.

        Putting together the considerations about row and column offsets, we can construct a formula $k = f(i, j)$ as follows:

        \begin{align*}
          f(i, j) =
          \begin{cases}
            j                                                            & \text{if } i = 1       \\
            \underbrace{3 + 3(i-2)}_{\text{previous i rows}} + j - i + 1 & \text{if } 2 \le i < n \\
          \end{cases}
        \end{align*}

        This simplifies to:

        \begin{align*}
          f(i, j) =
          \begin{cases}
            j          & \text{if } i = 1       \\
            2i + j - 2 & \text{if } 2 \le i < n \\
          \end{cases}
        \end{align*}

        Let's test this with a few cases from matrix $B_1$. We will set the non-zero values to their respective non-zero positions in the array $a$.

        \[
          B_1 =
          \begin{bmatrix}
            b_{11} & \boxed{b_{12}} & b_{13} & b_{14}         & b_{15}         \\
            b_{21} & b_{22}         & b_{23} & b_{24}         & b_{25}         \\
            b_{31} & \boxed{b_{32}} & b_{33} & b_{34}         & b_{35}         \\
            b_{41} & b_{42}         & b_{43} & \boxed{b_{44}} & b_{45}         \\
            b_{51} & b_{52}         & b_{53} & b_{54}         & \boxed{b_{55}} \\
          \end{bmatrix}
          =
          \begin{bmatrix}
            1 & \boxed{2} & 0 & 0          & 0          \\
            3 & 4         & 5 & 0          & 0          \\
            0 & \boxed{6} & 7 & 8          & 0          \\
            0 & 0         & 9 & \boxed{10} & 11         \\
            0 & 0         & 0 & 12         & \boxed{13}
          \end{bmatrix}
        \]

        \begin{align*}
          f(1,2) & = 2(1) + 2 - 2 \\
                 & = 2            \\
          f(3,2) & = 2(3) + 2 - 2 \\
                 & = 6            \\
          f(4,4) & = 2(4) + 4 - 2 \\
                 & = 10           \\
          f(5,5) & = 2(5) + 5 - 2 \\
                 & = 13           \\
        \end{align*}

        Based on these results, we can simplify the function further, since the case where $i = 1$ is handled correctly by the second case. Simplifying, we obtain a function for the 1-based index $k$ of the array $a$ that contains the non-zero elements of an $n \times n$ matrix:

        \begin{equation}
          f(i,j) = 2i + j - 2
        \end{equation}
        which holds $\forall i, j \in \mathbb{N} \mid 1 \le i \le n$ and $|j - i| \le 1$.

        Converting this function to work for a 0-based matrix of size $n \times n$:

        \[
          B_0 =
          \begin{bmatrix}
            b_{00} & \boxed{b_{01}} & b_{02} & b_{03}         & b_{04}         \\
            b_{10} & b_{11}         & b_{12} & b_{13}         & b_{14}         \\
            b_{20} & \boxed{b_{21}} & b_{22} & b_{23}         & b_{24}         \\
            b_{30} & b_{31}         & b_{32} & \boxed{b_{33}} & b_{34}         \\
            b_{40} & b_{41}         & b_{42} & b_{43}         & \boxed{b_{44}} \\
          \end{bmatrix}
          =
          \begin{bmatrix}
            1 & \boxed{2} & 0 & 0          & 0          \\
            3 & 4         & 5 & 0          & 0          \\
            0 & \boxed{6} & 7 & 8          & 0          \\
            0 & 0         & 9 & \boxed{10} & 11         \\
            0 & 0         & 0 & 12         & \boxed{13}
          \end{bmatrix}
        \]

        The index $k$ in an array of the non-zero elements of the matrix $B_0$ can be specified by adapting the function $f(i, j)$ to adding $2$ make up for the changed indices.
        \begin{align*}
          f_{0-\text{based}}(i,j) = 2(i) + j - 2 + 2
        \end{align*}

        This simplifies to:

        \begin{equation}
          f_{0-\text{based}}(i,j) = 2i + j
        \end{equation}

        $\forall i, j \in \mathbb{N} \mid 0 \le i \le n$ and $|j - i| \le 1$.

        We can validate this with several inputs:
        \begin{align*}
          f(0,1) & = 2(0) + 1 \\
                 & = 1        \\
          f(2,1) & = 2(2) + 1 \\
                 & = 5        \\
          f(3,3) & = 2(3) + 3 \\
                 & = 9        \\
          f(4,4) & = 2(4) + 4 \\
                 & = 12       \\
        \end{align*}
        Each of these values is the proper index $k$ for a 0-based array of the non-zero elements of the matrix in row-major order, as desired.

\end{enumerate}

\bigskip
\noindent\textbf{Problem 5:}  Consider two functions $f(n) = a \cdot n^2$, and $g(n) = b \cdot lg(n)$. ($lg(n)$ is $log$ base $2$ of $n$). For what value of $n$ do they intersect, and at which value(s) of the constants $a$ and $b$? Pick some values of $a$ and $b$ and then find $n$.  Experiment with different sets of values for $a$ and $b$. What trends do you observe? $a \cdot n^2$ and $b \cdot lg(n)$ are terms that might come from an expression representing the amount of work done by a piece of code. Looking for where they are equal will help you decide which part is dominant. We suggest solving this problem empirically, rather than mathematically, i.e. draw graph of the functions. Remember that $a$ and $b$ are constants, but may not necessarily be integers.

\noindent\textbf{Answer:}

\bigskip
\noindent\textbf{Problem 6:} What is the maximum size of a problem that can be solved in one hour if the algorithm takes $lg(n)$ microseconds?

Since the algorithm's speed is described in $lg(n)$, which is in base 2, the rest of our consideration will be in base 2. One microsecond is equal to $2^{-20}$ seconds. \textcite{null2023essentials}

\noindent\textbf{Answer:}

\bigskip
\noindent\textbf{Problem 7:} What is the maximum size of a problem that can be solved in one hour if the algorithm take $n^3$ microseconds?

As above, this analysis will be done according to base 2. One microsecond is equal to $2^{-20}$ seconds. \textcite{null2023essentials}

\noindent\textbf{Answer:}

\pagebreak

\section*{Source Code}
The full source code, including LaTeX, tooling, other artifacts, and git history is available on my Github repository for the class here:
\begin{center}
  \url{https://github.com/jllovet/jhu-en-605-202-su26-data-structures}
\end{center}

\printbibliography

\end{document}